---
title: "Joel_IVRS_Script"
author: "Harun"
date: 'June 6, 2022'
output: pdf_document
font: Calibri Light
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Libraries
library(ggplot2)
library(dplyr)

# Read in data
IVRS <- read.csv("Data/Closed_Iowa_Vocational_Rehabilitation_Cases.csv")
IDB <- read.csv("Data/Successful_Employment_for_Blind_Iowans_by_Federal_Fiscal_Year.csv")
```

# Exploration of the Iowa Vocational Rehabilitation Services (IVRS) Data

This data comes from the **[data.iowa.gov](https://https://data.iowa.gov/Vocational-Rehabilitation/Closed-Iowa-Vocational-Rehabilitation-Cases/gcya-t3y9)** website and provides information on closed cases where the individual received services from IVRS. This data includes cases closed after October 1, 2008.

## Number of Cases by Gender

This section looks at the number of cases for each gender. 

From this graph we can see that there were more cases for men although it can be noted that the number of cases for women isn't to far off from the total of men's cases.

```{r Number of Cases by Gender, echo=FALSE}
# get number of cases by gender 
gender.cases <- IVRS %>%
  count(Gender, name = "Total.Cases") %>%
  group_by(Gender)

# change factors
gender.cases$Gender <- factor(gender.cases$Gender, levels=c("MALE", "FEMALE", "DID NOT SELF-IDENTIFY"))

#plot data
ggplot(data = gender.cases, aes(x = Gender, y = Total.Cases)) +
  geom_bar(fill = "orange", stat = "identity") +
  geom_text(aes(label=Total.Cases), vjust=-0.5) +
  xlab("Gender") +
  ylab("Number of Cases")
```

## Number of Cases by Age

### Age at Start of Application

This section looks at the number of cases by the age of the client when the case was started.

Initially I put the data into a bar graph but it is difficult to read the various ages and the bar labels are crowded. 

```{r Number of Cases by Age at Start, echo=FALSE}
# get number of cases by age at start of application 
age.start.cases <- IVRS %>%
  count(Age.at.Application, name = "Total.Cases") %>%
  group_by(Age.at.Application)

# plot data
ggplot(data = age.start.cases, aes(x = as.factor(Age.at.Application), y = Total.Cases)) +
  geom_bar(fill = "orange", stat = "identity") +
  geom_text(aes(label=Total.Cases), vjust=-0.5) +
  xlab("Age at Start of Application") +
  ylab("Number of Cases")
```

I then looked at a table of the data sorted by the number of cases which was more helpful.

From this table we can see that the majority of cases were from those in their mid teen to mid twenties (as could be seen in the bar graph). 

```{r, echo=FALSE}
# create descending table
age.start.cases %>% arrange(desc(Total.Cases))
```

### Age at Closure of Application

This section looks at the number of cases by the age of the client when the case was closed.

As before I put the data into a bar graph but again it is difficult to read the various ages and the bar labels are crowded. We do however see the same trend as before of the majority of those with closed cases are in their mid teen to mid twenties.



```{r Number of Cases by Age at Closure, echo=FALSE}
# get number of cases by age at closure of case
age.closure.cases <- IVRS %>%
  count(Age.at.Closure, name = "Total.Cases") %>%
  group_by(Age.at.Closure)
  
# plot data
ggplot(data = age.closure.cases, aes(x = Age.at.Closure, y = Total.Cases)) +
  geom_bar(fill = "orange", stat = "identity") +
  xlab("Age at Closure of Application") +
  ylab("Number of Cases")
```

I then looked at a table of the data sorted by the number of cases.

From this table we can see that the majority of cases were actually from those in their late teen to late twenties.

```{r, echo=FALSE}
# create descending table 
age.closure.cases %>% arrange(desc(Total.Cases))
```

### Age Span of Clients

I then looked at the age span of the cases by subtracting the persons age at the end of their case from their age at the start of the case.

We can see that the majority of the cases were only open for 1 - 3 years. This makes sense since from our previous graphs it appeared that those in the majority age range for their case start age were in the majority for their case closed age.

```{r Age Spans, echo=FALSE}
# get number of years between start age and age at closure
age.span <- transform(IVRS, Age.Span = Age.at.Closure - Age.at.Application)

# get number of cases for each age span
age.span <- age.span %>%
  count(Age.Span, name = "Total") %>%
  group_by(Age.Span)
  
# plot data
ggplot(data = age.span, aes(x = as.factor(Age.Span), y = Total)) +
  geom_bar(fill = "orange", stat = "identity") +
  geom_text(aes(label=Total), vjust=-0.5) +
  xlab("Age Span") +
  ylab("Number of Cases")
```

## Minority Groups

This section looks at the minority groups. 

We can see that most of the cases were non-minority clients.

```{r Minority Groups, echo=FALSE}
# get number of cases by minority group 
minority.group.cases <- IVRS %>%
  count(Minority.Group, name = "Total.Cases") %>%
  group_by(Minority.Group)

# change factors
minority.group.cases$Minority.Group <- factor(minority.group.cases$Minority.Group, levels=c("Minority", "Non-Minority", "Did Not Self-identify", "NULL"))

  
# plot data
ggplot(data = minority.group.cases, aes(x = Minority.Group, y = Total.Cases)) +
  geom_bar(fill = "orange", stat = "identity") +
  xlab("Minority Group") +
  ylab("Number of Cases")
```

## Disability Groups and Types

This section looks at the disability groups and types. Please note that SD stands for Significantly Disabled and MSD stands for Most Significantly Disabled.

### Disability Group

We can see that most of the cases were those who were SD were higher than those who were MSD.

```{r Disability Groups, echo=FALSE}
# get number of cases by disability group 
disability.groups <- IVRS %>%
  count(Disability.Category, name = "Total.Cases") %>%
  group_by(Disability.Category)

# change factors
disability.groups$Disability.Category <- factor(disability.groups$Disability.Category, levels=c("MSD", "SD", "Other"))

# plot data
ggplot(data = disability.groups, aes(x = Disability.Category, y = Total.Cases)) +
  geom_bar(fill = "orange", stat = "identity") +
  xlab("Disability Category") +
  ylab("Number of Cases")
```
### Disability Type

This section looks at disability by type.

We can see that those with a cognitive disability had the most number of closed cases with those having a psychosocial mental disability being second.

```{r Disability Type, echo=FALSE}
# get number of cases by disability type 
disability.types <- IVRS %>%
  count(Disability.Type, name = "Total.Cases") %>%
  group_by(Disability.Type)

# factor
disability.types$Disability.Type <- factor(disability.types$Disability.Type,
                                           levels = disability.types$Disability.Type[order(disability.types$Total.Cases, decreasing = TRUE)])

# plot data
ggplot(data = disability.types, aes(x = Disability.Type, y = Total.Cases)) +
  geom_bar(fill = "orange", stat = "identity") +
  xlab("Disability Type") +
  ylab("Number of Cases") +
  theme(axis.text.x = element_text(angle = 9))
```

### Disability by Group and Type

This section looks at both group and type of disability.

We can see that those with a cognitive disability had the most number of closed cases among all groups and those having a psychosocial mental disability being second as was found in the last graph.

```{r Disability Category by Type, echo=FALSE}
# get number of cases by disability type 
disability.category.types <- IVRS %>%
  group_by(Disability.Category, Disability.Type) %>%
  count(Disability.Type, name = "Total")

# change factors
disability.category.types$Disability.Category <- factor(disability.category.types$Disability.Category, levels=c("MSD", "SD", "Other"))

# plot data
ggplot(data = disability.category.types, aes(x = Disability.Category, 
                                    y = Total, 
                                    fill = Disability.Type)) +
  geom_bar(stat = "identity", position = "dodge") +
  xlab("Disability Category") +
  ylab("Number of Cases")
```

## Monthly Cases

This section looks at how long the closed cases were opened. 

I created a bar chart to view this data, however there are a large number of data points and it is difficult to see specifics. We can see that the majority of cases were opened between about 10 - 50 months (about 1 - 5 years).

```{r Monthly Cases, echo=FALSE}
# get number of cases by months opened 
months.opened <- IVRS %>%
  count(Months.Case.Open, name = "Total.Cases") %>%
  group_by(Months.Case.Open)
  
# plot data
ggplot(data = months.opened, aes(x = Months.Case.Open, y = Total.Cases)) +
  geom_bar(fill = "orange", stat = "identity") +
  xlab("Number of Months Opened") +
  ylab("Number of Cases")
```

Creating a sorted table shows that most of the cases were opened for about 1 $\frac{1}{2}$ - 2 years.

```{r, echo=FALSE}
# create descending table
months.opened %>% arrange(desc(Total.Cases))
```

## Yearly Cases

This section looks at how many cases were closed each year. 

We can see that the number of closed cases has remained relatively the same over the course of the data sets history.

```{r Yearly Cases, echo=FALSE}
# get number of cases by year 
yearly.cases <- IVRS %>%
  count(Federal.Fiscal.Year, name = "Total.Cases") %>%
  group_by(Federal.Fiscal.Year)
  
# plot data
ggplot(data = yearly.cases, aes(x = as.factor(Federal.Fiscal.Year), y = Total.Cases)) +
  geom_bar(fill = "orange", stat = "identity") +
  geom_text(aes(label=Total.Cases), vjust=-0.5) +
  xlab("Year") +
  ylab("Number of Cases")
```

## Date Span

This section looks at the span of the cases by start date of application to date of closure.

A bar graph is somewhat helpful however it is difficult to see specifics. It appears that the majority of the cases are opened between 0 - 1000 days (about 0 - 32 months or about 0 - 3 years)

```{r Date Span, echo=FALSE}
# get number of cases by date span 
date.span <- IVRS %>%
  count(date.span = (as.Date(Closure.Date,
                format = '%m/%d/%Y') -
          as.Date(Application.Date,
                  format = '%m/%d/%Y')), 
        name = "Total.Cases") %>%
  group_by(date.span)

# plot data
ggplot(data = date.span, aes(x = date.span, y = Total.Cases)) +
  geom_bar(fill = "orange", stat = "identity") +
  xlab("Number of Days Case was Opened") +
  ylab("Number of Cases")
```

Looking at a table we can see the highest number of total cases is 462 days which is about 1 $\frac{1}{4}$ years which is close to the range we found from looking at the months a case was open. 

```{r, echo=FALSE}
# create descending table
date.span %>% arrange(desc(Total.Cases))
```

## Cases by Client County

This section looks at the clients for each case by county they are from. 

Initially I attempted to make a bar chart hoping it would hold all of Iowa's 99 counties. However I learned that there are many counties in the data set outside of Iowa which made it almost impossible to view the data in a bar graph format.

```{r Cases by Client County, echo=FALSE}
# get number of cases by clients county 
county.cases <- IVRS %>%
  count(Client.County, name = "Total.Cases") %>%
  group_by(Client.County)
  
# plot data
ggplot(data = county.cases, aes(x = Client.County, y = Total.Cases)) +
  geom_bar(fill = "orange", stat = "identity") +
  xlab("County") +
  ylab("Number of Cases") +
  theme(axis.text.x = element_text(angle = 90))
```

I then opted to make a table showing the number of closed cases for each county sorted by those with the highest number of cases. 

From this table we can see that Polk and Linn counties have the most closed cases. A quick **[Google search](https://www.iowa-demographics.com/counties_by_population)** showed from the 2020 American Community Survey (ACS) data that the counties with the highest population were Polk and Linn so this data aligns with what we would expect.
```{r, echo=FALSE}
# create descending table
county.cases %>% arrange(desc(Total.Cases))
```


# Cases by Office Area

This section looks at the number of closed cases by the office area.

We can see that the Des Moines (026) office  has the most number of closed cases.

```{r Cases by Office Area, echo=FALSE}
# get number of cases by office area 
office.cases <- IVRS %>%
  count(Office.Area, name = "Total.Cases") %>%
  group_by(Office.Area)
  
# plot data
ggplot(data = office.cases, aes(x = reorder(Office.Area, -Total.Cases), y = Total.Cases)) +
  geom_bar(fill = "orange", stat = "identity") +
  geom_text(aes(label=Total.Cases), vjust=0) +
  xlab("Office") +
  ylab("Number of Cases") +
  theme(axis.text.x = element_text(angle = 30))
```


