---
title: "Xingrui_ACS5_AgeRace"
author: "Xingrui Xue"
date: '2022-06-17'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidycensus)
library(tidyverse)
library(openxlsx)
library(stringr)

acs5 <- load_variables(2020, "acs5")
```

Population of Iowans 5-64 years of age with disabilities (Disaggregate by age groups)
```{r}
DisabledHearing <- c("5 to 17 years" = c("B18102_007", "B18102_026"), "18 to 34 years" = c("B18102_010", "B18102_029"), "35 to 64 years" = c("B18102_013", "B18102_032"))
DisabledVision <- c("5 to 17 years" = c("B18103_007", "B18103_026"), "18 to 34 years" = c("B18103_010", "B18103_029"), "35 to 64 years" = c("B18103_007", "B18103_032"))
DisabledCognitive <- c("5 to 17 years" = c("B18104_004", "B18104_020"), "18 to 34 years" = c("B18104_007", "B18104_023"), "35 to 64 years" = c("B18104_010", "B18104_026"))
DisabledAmbulatory <- c("5 to 17 years" = c("B18105_004", "B18105_020"), "18 to 34 years" = c("B18105_007", "B18105_023"), "35 to 64 years" = c("B18105_010", "B18105_026"))
DisabledSelfCare <- c("5 to 17 years" = c("B18106_004", "B18106_020"), "18 to 34 years" = c("B18106_007", "B18106_023"), "35 to 64 years" = c("B18106_010", "B18106_026"))
DisabledIndependentLiving <- c("5 to 17 years" = c("B18107_004", "B18107_020"), "18 to 34 years" = c("B18107_007", "B18107_023"), "35 to 64 years" = c("B18107_010", "B18107_026"))

DisabledType <- get_acs(geography = "county",
                        variables = c(Hearing = DisabledHearing, Vision = DisabledVision, Cognitive = DisabledCognitive, "Self-care" = DisabledSelfCare, "Independent living" = DisabledIndependentLiving),
                        year = 2020,
                        state = "IA")

DisabledType$variable <- gsub("(years).*", "\\1", DisabledType$variable)

DisabledType <- DisabledType %>%
  group_by(NAME, variable) %>%
  summarise(estimate = sum(estimate),
            moe = sum(moe))

DisabledType_pivot <- DisabledType %>% separate(variable, into = c("Type", "Age"), sep = "\\.") %>%
  pivot_wider(names_from = Type, values_from = c(estimate, moe))

DisabledType_sum <- aggregate(estimate ~ NAME, data = DisabledType, sum) %>%
  slice(rep(1:n(), each = 3))

DisabledType_pivot <- DisabledType_pivot %>%
  ungroup() %>%
  mutate(Percentage = rowSums(DisabledType_pivot[3:7] / DisabledType_sum$estimate) * 100)

names(DisabledType_pivot) <- gsub("estimate_", "", names(DisabledType_pivot))
DisabledType_pivot$NAME <- gsub(" County, Iowa", "", DisabledType_pivot$NAME)
names(DisabledType_pivot)[names(DisabledType_pivot) == "NAME"] <- "County"
```

Testing the reliability of DisabledType_pivot data
```{r}
CV_Cognitive <- ((DisabledType_pivot$moe_Cognitive / 1.645) / DisabledType_pivot$Cognitive) * 100
CV_Hearing <- ((DisabledType_pivot$moe_Hearing / 1.645) / DisabledType_pivot$Hearing) * 100
CV_IndependentLiving  <- ((DisabledType_pivot$`moe_Independent living` / 1.645) / DisabledType_pivot$`Independent living`) * 100
CV_Selfcare <- ((DisabledType_pivot$`moe_Self-care` / 1.645) / DisabledType_pivot$`Self-care`) * 100
CV_Vision <- ((DisabledType_pivot$moe_Vision / 1.645) / DisabledType_pivot$Vision) * 100

DisabledType_pivot$CV_Cognitive <- ifelse(CV_Cognitive > 30, "Low reliability",
                                          ifelse(CV_Cognitive <= 15, "High reliability",
                                                 "Median reliability"))
DisabledType_pivot$CV_Hearing <- ifelse(CV_Hearing > 30, "Low reliability",
                                        ifelse(CV_Hearing <= 15, "High reliability",
                                               "Median reliability"))
DisabledType_pivot$CV_IndependentLiving <- ifelse(CV_IndependentLiving > 30, "Low reliability",
                                                  ifelse(CV_IndependentLiving <= 15, "High reliability",
                                                         "Median reliability"))
DisabledType_pivot$CV_Selfcare <- ifelse(CV_Selfcare > 30, "Low reliability",
                                         ifelse(CV_Selfcare <= 15, "High reliability",
                                                "Median reliability"))
DisabledType_pivot$CV_Vision <- ifelse(CV_Vision > 30, "Low reliability",
                                       ifelse(CV_Vision <= 15, "High reliability",
                                              "Median reliability"))

View(DisabledType_pivot)
write.xlsx(DisabledType_pivot, "DisabledType.xlsx")
```

Population of Iowans 5-64 years of age with disabilities (Disaggregate by difficulty type)
```{r}
DisabledType_pivot1 <- DisabledType %>% separate(variable, into = c("Type", "Age"), sep = "\\.") %>%
  pivot_wider(names_from = Age, values_from = c(estimate, moe))

DisabledType_sum1 <- aggregate(estimate ~ NAME, data = DisabledType, sum) %>%
  slice(rep(1:n(), each = 5))

DisabledType_pivot1 <- DisabledType_pivot1 %>%
  ungroup() %>%
  mutate(Percentage = rowSums(DisabledType_pivot1[3:5] / DisabledType_sum1$estimate) * 100)

names(DisabledType_pivot1) <- gsub("estimate_", "", names(DisabledType_pivot1))
DisabledType_pivot1$NAME <- gsub(" County, Iowa", "", DisabledType_pivot1$NAME)
names(DisabledType_pivot1)[names(DisabledType_pivot1) == "NAME"] <- "County"

write.xlsx(DisabledType_pivot1, "DisabledType1.xlsx")
```

Race and ethnic breakdown of populations for individuals with disabilities (Disaggregate by age groups)
```{r}
DisabledWhite <- c("Under 18 years" = "B18101A_003", "18 to 64 years" = "B18101A_006", "65 years and over" = "B18101A_009")
DisabledBlack <- c("Under 18 years" = "B18101B_003", "18 to 64 years" = "B18101B_006", "65 years and over" = "B18101B_009")
DisabledIndianAlaska <- c("Under 18 years" = "B18101C_003", "18 to 64 years" = "B18101C_006", "65 years and over" = "B18101C_009")
DisabledAsian <- c("Under 18 years" = "B18101D_003", "18 to 64 years" = "B18101D_006", "65 years and over" = "B18101D_009")
DisabledHawaiianPacific <- c("Under 18 years" = "B18101E_003", "18 to 64 years" = "B18101E_006", "65 years and over" = "B18101E_009")
DisabledOtherAlone <- c("Under 18 years" = "B18101F_003", "18 to 64 years" = "B18101F_006", "65 years and over" = "B18101F_009")
DisabledTwoMore <- c("Under 18 years" = "B18101G_003", "18 to 64 years" = "B18101G_006", "65 years and over" = "B18101G_009")
DisabledWhiteAlone <- c("Under 18 years" = "B18101H_003", "18 to 64 years" = "B18101H_006", "65 years and over" = "B18101H_009")
DisabledHispanicLatino <- c("Under 18 years" = "B18101I_003", "18 to 64 years" = "B18101I_006", "65 years and over" = "B18101I_009")

DisabledRace <- get_acs(geography = "county",
                        variables = c("WHITE ALONE" = DisabledWhite, "BLACK OR AFRICAN AMERICAN ALONE" = DisabledBlack, "AMERICAN, INDIAN AND ALASKA NATIVE ALONE" = DisabledIndianAlaska, "ASIAN ALONE" = DisabledAsian, "NATIVE HAWAIIAN AND OTHER PACIFIC ISLANDER ALONE" = DisabledHawaiianPacific, "SOME OTHER RACE ALONE" = DisabledOtherAlone, "TWO OR MORE RACES" = DisabledTwoMore, "WHITE ALONE, NOT HISPANIC OR LATINO" = DisabledWhiteAlone, "HISPANIC OR LATINO" = DisabledHispanicLatino),
                        year = 2020,
                        state = "IA")

DisabledRace_pivot <- DisabledRace %>% separate(variable, into = c("Race", "Age"), sep = "\\.") %>%
  pivot_wider(names_from = Race, values_from = c(estimate, moe))

DisabledRace_sum <- aggregate(estimate ~ NAME, data = DisabledRace, sum) %>%
  slice(rep(1:n(), each = 3))

DisabledRace_pivot <- DisabledRace_pivot %>%
  mutate(Percentage = rowSums(DisabledRace_pivot[4:12] / DisabledRace_sum$estimate) * 100)

names(DisabledRace_pivot) <- gsub("estimate_", "", names(DisabledRace_pivot))
DisabledRace_pivot$NAME <- gsub(" County, Iowa", "", DisabledRace_pivot$NAME)
names(DisabledRace_pivot)[names(DisabledRace_pivot) == "NAME"] <- "County"

write.xlsx(DisabledRace_pivot, "DisabledRace.xlsx")
```

Population of Iowans 18-64 years of age with disabilities and without disabilities
```{r}
Disabled18to34 <- c(Total = c("B18101_009", "B18101_028"), Disabled = c("B18101_010", "B18101_029"))
Nondisabled18to34 <- c(Total = c("B18101_009", "B18101_028"), Nondisabled = c("B18101_011", "B18101_030"))
Disabled35to64 <- c(Total = c("B18101_012", "B18101_031"), Disabled = c("B18101_013", "B18101_032"))
Nondisabled35to64 <- c(Total = c("B18101_012", "B18101_031"), Nondisabled = c("B18101_014", "B18101_033"))

population <- get_acs(geography = "county",
                      variables = c("18 to 34 years" = c(Disabled18to34, Nondisabled18to34), "35 to 64 years" = c(Disabled35to64, Nondisabled35to64)),
                      year = 2020,
                      state = "IA")

population$variable <- str_sub(population$variable, end = -2)

population <- population %>%
  group_by(NAME, variable) %>%
  summarise(estimate = sum(estimate),
            moe = sum(moe)) %>%
  pivot_wider(names_from = variable,
              values_from = c(estimate, moe))

names(population) <- gsub("estimate_", "", names(population))
population$NAME <- gsub(" County, Iowa", "", population$NAME)
names(population)[names(population) == "NAME"] <- "County"
```

Testing the reliability of population data
```{r}
CV_18_to_34_Disabled <- ((population$`moe_18 to 34 years.Disabled` / 1.645) / population$`18 to 34 years.Disabled`) * 100
CV_18_to_34_Nondisabled <- ((population$`moe_18 to 34 years.Nondisabled` / 1.645) / population$`18 to 34 years.Nondisabled`) * 100
CV_35_to_64_Disabled <- ((population$`moe_35 to 64 years.Disabled` / 1.645) / population$`35 to 64 years.Disabled`) * 100
CV_35_to_64_Nondisabled <- ((population$`moe_35 to 64 years.Nondisabled` / 1.645) / population$`35 to 64 years.Nondisabled`) * 100

population$CV_18_to_34_Disabled <- ifelse(CV_18_to_34_Disabled > 30, "Low reliability",
                                          ifelse(CV_18_to_34_Disabled <= 15, "High reliability",
                                                 "Median reliability"))
population$CV_18_to_34_Nondisabled <- ifelse(CV_18_to_34_Nondisabled > 30, "Low reliability",
                                          ifelse(CV_18_to_34_Nondisabled <= 15, "High reliability",
                                                 "Median reliability"))
population$CV_35_to_64_Disabled <- ifelse(CV_35_to_64_Disabled > 30, "Low reliability",
                                          ifelse(CV_35_to_64_Disabled <= 15, "High reliability",
                                                 "Median reliability"))
population$CV_35_to_64_Nondisabled <- ifelse(CV_35_to_64_Nondisabled > 30, "Low reliability",
                                          ifelse(CV_35_to_64_Nondisabled <= 15, "High reliability",
                                                 "Median reliability"))

View(population)
write.xlsx(population, "Population.xlsx")
```

Race and ethnic breakdown of populations (18 to 64 years) for both individuals with disabilities and without disabilities
```{r}
White <- c("Total White"= "B18101A_005", "Disabled White" = "B18101A_006", "Nondisabled White" = "B18101A_007")
Black <- c("Total Black" ="B18101B_005", "Disabled Black" = "B18101B_006", "Nondisabled Black" = "B18101B_007")
IndianAlaska <- c("Total American Indian and Alaska Native" = "B18101C_005", "Disabled American Indian and Alaska Native" = "B18101C_006", "Nondisabled American Indian and Alaska Native" = "B18101C_007")
Asian <- c("Total Asian" = "B18101D_005", "Disabled Asian" = "B18101D_006", "Nondisabled Asian" = "B18101D_007")
HawaiianPacific <- c("Total Native Hawaiian and Other Pacific Islander" = "B18101E_005", "Disabled Native Hawaiian and Other Pacific Islander" = "B18101E_006", "Nondisabled Native Hawaiian and Other Pacific Islander" = "B18101E_007")
OtherAlone <- c("Total Some Other Race Alone" = "B18101F_005", "Disabled Some Other Race Alone" = "B18101F_006", "Nondisabled Some Other Race Alone" = "B18101F_007")
TwoMore <- c("Total Two Or More Races" = "B18101G_005", "Disabled Two Or More Races" = "B18101G_006", "Nondisabled Two Or More Races" = "B18101G_007")
WhiteAlone <- c("Total White Alone" = "B18101H_005", "Disabled White Alone" = "B18101H_006", "Nondisabled White Alone" = "B18101H_007")
HispanicLatino <- c("Total Hispanic Or Latino" = "B18101I_005", "Disabled Hispanic Or Latino" = "B18101I_006", "Nondisabled Hispanic Or Latino" = "B18101I_007")

RaceEthnic <- get_acs(geography = "county",
                      variables = c(White, Black, IndianAlaska, Asian, HawaiianPacific, OtherAlone, TwoMore, WhiteAlone, HispanicLatino),
                      year = 2020,
                      state = "IA")

RaceEthnic <- RaceEthnic %>%
  pivot_wider(names_from = variable,
              values_from = c(estimate, moe))

names(RaceEthnic) <- gsub("estimate_", "", names(RaceEthnic))
RaceEthnic$NAME <- gsub(" County, Iowa", "", RaceEthnic$NAME)
names(RaceEthnic)[names(RaceEthnic) == "NAME"] <- "County"
```

Testing the reliability of RaceEthnic data
```{r}
CV_White_Disabled <- ((RaceEthnic$`moe_Disabled White` / 1.645) / RaceEthnic$`Disabled White`) * 100
CV_White_Nondisabled <- ((RaceEthnic$`moe_Nondisabled White` / 1.645) / RaceEthnic$`Nondisabled White`) * 100
CV_Black_Disabled <- ((RaceEthnic$`moe_Disabled Black` / 1.645) / RaceEthnic$`Disabled Black`) * 100
CV_Black_Nondisabled <- ((RaceEthnic$`moe_Nondisabled Black` / 1.645) / RaceEthnic$`Nondisabled Black`) * 100
CV_IndianAlaska_Disabled <- ((RaceEthnic$`moe_Disabled American Indian and Alaska Native` / 1.645) / RaceEthnic$`Disabled American Indian and Alaska Native`) * 100
CV_IndianAlaska_Nondisabled <- ((RaceEthnic$`moe_Nondisabled American Indian and Alaska Native` / 1.645) / RaceEthnic$`Nondisabled American Indian and Alaska Native`) * 100
CV_Asian_Disabled <- ((RaceEthnic$`moe_Disabled Asian` / 1.645) / RaceEthnic$`Disabled Asian`) * 100
CV_Asian_Nondisabled <- ((RaceEthnic$`moe_Nondisabled Asian` / 1.645) / RaceEthnic$`Nondisabled Asian`) * 100
CV_HawaiianPacific_Disabled <- ((RaceEthnic$`moe_Disabled Native Hawaiian and Other Pacific Islander` / 1.645) / RaceEthnic$`Disabled Native Hawaiian and Other Pacific Islander`) * 100
CV_HawaiianPacific_Nondisabled <- ((RaceEthnic$`moe_Nondisabled Native Hawaiian and Other Pacific Islander` / 1.645) / RaceEthnic$`Disabled Native Hawaiian and Other Pacific Islander`) * 100
CV_OtherAlone_Disabled <- ((RaceEthnic$`moe_Disabled Some Other Race Alone` / 1.645) / RaceEthnic$`Disabled Some Other Race Alone`) * 100
CV_OtherAlone_Nondisabled <- ((RaceEthnic$`moe_Nondisabled Some Other Race Alone` / 1.645) / RaceEthnic$`Nondisabled Some Other Race Alone`) * 100
CV_TwoMore_Disabled <- ((RaceEthnic$`moe_Disabled Two Or More Races` / 1.645) / RaceEthnic$`Disabled Two Or More Races`) * 100
CV_TwoMore_Nondisabled <- ((RaceEthnic$`moe_Nondisabled Two Or More Races` / 1.645) / RaceEthnic$`Nondisabled Two Or More Races`) * 100
CV_WhiteAlone_Disabled <- ((RaceEthnic$`moe_Disabled White Alone` / 1.645) / RaceEthnic$`Disabled White Alone`) * 100
CV_WhiteAlone_Nondisabled <- ((RaceEthnic$`moe_Nondisabled White Alone` / 1.645) / RaceEthnic$`Nondisabled White Alone`) * 100
CV_HispanicLatino_Disabled <- ((RaceEthnic$`moe_Disabled Hispanic Or Latino` / 1.645) / RaceEthnic$`Disabled Hispanic Or Latino`) * 100
CV_HispanicLatino_Nondisabled <- ((RaceEthnic$`moe_Nondisabled Hispanic Or Latino` / 1.645) / RaceEthnic$`Nondisabled Hispanic Or Latino`) * 100
```

Earnings of disabled population.
```{r}
earning <- get_acs(geography = "county",
                   variables = "B18140_002",
                   year = 2020,
                   state = "IA")

names(earning) <- gsub("estimate_", "", names(earning))
earning$NAME <- gsub(" County, Iowa", "", earning$NAME)
names(earning)[names(earning) == "NAME"] <- "County"

write.xlsx(earning, "EarningDisabled.xlsx")
```

